\fancychapter{Introduction}
 With the evolution of social network websites like Facebook and Twitter, the amount of pertinent content about a specif issue is increasing dramatically, which calls for new ways to make sense and catalog this data.
The usage of social networks for branding quality and on-line marketing is specially compelling since 19\% of all tweets ~\cite{Jansen2009} and 32\% ~\cite{Melville2009} of blog posts are about brands or products.
On the other hand, finding topic sensitive information on social networks is extremely complicate due to the fact that documents have very little content, slang vocabulary and orthographically mistakes or abbreviations.

~\citet{Asur2010} successfully predicted box-office revenues by monitoring the rate of creation of new topics based on debuting movies. ~\citet{Asur2010} was able to outperformed some market-based predictors.

Academic and enterprise world is now starting to look at Machine Learning for new ways to achieve revenue and visualize data representing the way the world works. 
As a consequence, the Machine Learning course at Standford is the one with more students enrolling this year \footnote{http://www.forbes.com/sites/anthonykosner/2013/12/29/why-is-machine-learning-cs-229-the-most-popular-course-at-stanford/} with more than 760 students enrolled.

 ~\cite{Le2011} was able to achieved 81.7 percent accuracy in detecting human faces, 76.7 percent accuracy when identifying human body parts and 74.8 percent accuracy when identifying cats. He used a 9-layered locally connected sparse auto-encoder with pooling and local contrast normalization on a large dataset of images (the model has 1 billion connections, the dataset has 10 million 200x200 pixel images downloaded from the Internet) trained using model parallelism and asynchronous SGD on a cluster with 1,000 machines (16,000 cores) during three days. Even though the amount of computing power used in this project was of several order of magnitude, it is remarkable how an unsupervised algorithm could achieve such results.

Even though a lot of solutions have arisen in order to automate real time searches, topic categorization and many other data intensive tasks, Twitter still uses humans in order to deliver ads to trending queries states Edwin Chen's Data Scientist Responsible for ads quality at Twitter. On his blog post \footnote{http://blog.echen.me/2013/01/08/improving-twitter-search-with-real-time-human-computation/} Edwin describes the process of Twitter to deliver real time adds to trending queries, the main problems that arise in the Twitter platform in order to identify rising topics are mainly:
\begin{itemize}
  \item The queries people perform have never before been seen, so it's impossible to know beforehand what they mean.
  \item Since the spikes in search queries are short-lived, there's only a short window of opportunity to learn what they mean.
\end{itemize}
This means that when an event happen, people immediately come to Twitter in order to know what is happening in a determined place in real time. Twitter solves this issue by monitoring which queries are currently popular in real time, using a Storm topology \footnote{http://storm-project.net/} and after the queries are identified, they are sent to a Thrift API \footnote{http://thrift.apache.org/} that dispatches the query to Amazon's Mechanical Turk \footnote{https://www.mturk.com/mturk/} service where real people will be asked a variety of questions about the query.

Social Media Analytics is another raising topic which draws from Social Network Analysis, Machine Learning, Data Mining, Information Retrieval (IR), and Natural Language Processing (NLP). As stated by Melville et al. in ~\cite{Melville2009} 32\% of the 200 million bloggers world wide blog about opinions on products and brands, 71\% of the 625 million active Internet users actually read blogs and more importantly that 78\% of respondents put their trust in the opinion of other consumers where only only 57\% of consumers trust advertising in traditional media and even worst only 34\% of consumers put their trust in such advertising. This kind of data drives companies to Social Media Analytics in a way to know what people are saying on the web about their companies and products. This new worry has brought to life a lot of new startups like Sumal\footnote{https://sumall.com/} or ThoughtBuzz\footnote{http://www.thoughtbuzz.net/} but also solutions from the old players like IBM \footnote{http://www-01.ibm.com/software/analytics/solutions/customer-analytics/social-media-analytics/} and SAS \footnote{http://www.sas.com/software/customer-intelligence/social-media-analytics.html}

Its also important to notice that in the last few years Data Science/Analysis has been a trending topic, mostly due to the fact that big dot-com companies have been making lots of money through exploiting user specific information in order to deliver ads and sell products. No wonder that if you look that in the top 10 ebooks sold by O'Reilly throughout 2013, four are about data science \footnote{http://shop.oreilly.com/category/deals/best\-of\-oreilly\-dotd.do?code=DEAL\&cmp=tw\-na\-books\-videos\-info\-authornote\_best\_of\_2013}.

In this project we will focus on using an unsupervised learning technique based on neural networks named Self-Organizing Maps ~\cite{Kohonen1990} in order to detect topics in Twitter posts, by using the Social Network users as base neurons for clustering. After the network is trained, it will be possible to categorize tweets in real time. This approach will be better described in subsection ~\ref{sub:objectives}.

This document is organized as follows. First will be dedicated to explain some basics concepts like Document CLustering and specifically Self-Organizing Maps in Section ~\ref{sec:basic_concepts}.  Further in, Section ~\ref{sec:related_work} will be dedicated to the state of the art solutions related not only to topic detection but also to twitter data analysis and Self-Organizing Maps.In Section ~\ref{sec:architecture} Architecture of the purposed solution and  at Section ~\ref{sec:evaluation_metrics} it will be discussed how to evaluate results achieved. Finally we will this report by referencing some possible future work and with a brief conclusion at Section ~\ref{sec:coclusions_future_work}.
 
%TODO Check if Im Answering these questions
\begin{enumerate}
 \item what is the background of your work?
 \item how do your work fit in todays knowledge? Have you done something that does not exist in the market? What are the differences? What is missing in nowadays products and solutions?
 \item is your thesis made as part of a larger project? If so, describe it.
 \item is your thesis usefull for some work environment? If so, describe it.
\end{enumerate}

\section{Motivation}

%TODO I dont have motivation
Given the general description provided previously, what is the motivation of your work. Explain why the product or solution developed in the course of your thesis is important.

\section{Objectives}

%TODO Change text to what Ive actually done
The objective of this project is clear: finding topics on Tweets by analyzing their textual specific characteristics, like number of characters in a tweet, hashtags, is a retweeted \footnote{retweet is when a user shares a tweet that is not his}, etc., and contextualize the social network involving the person that did the tweet by retrieving user specific profile information.

After building the dataset, it will be needed to train the Self-Organizing map neural network in order to be able to have the network ready for clustering classification of each future tweet that will be added to it. After the SOM training it will be necessary to categorize the clusters in order to know which topic they belong to. When this step ha ended it will be possibly to get new tweets categorized on the moment they enter the network without further delay.

Finally it will be presented a website where a user will be able to login with his twitter account and see all of his tweets being categorization in the moment they are clustered. After all the user tweets are clustered there will a graphical presentation of the user twitter usage where it will be possible to see multiple statistical information such as the the topics a user is more interested in and his own twitter Self-Organizing map network of topics with his friends, the wire-frames of the website can be seen in the attachments Section ~\ref{sec:attachments} in Figure ~\ref{fig:wireframes}.
 

\section{Main contributions}
%TODO Explain all the open source lybraries that Ive published
Do not forget this one. Notice that the objectives is what you have proposed to do. Main contributions are the innovations of your work, or, in other cases, what your work is really good at. If you submitted/published an article in a peer-reviewed conference or journal, do not forget to state here.

\section{Dissertation outline}
%TODO Write the outline
Explain how did you organized your thesis.

\cleardoublepage 
